\artus -- Framework for Event-based Data Analysis in High Energy Physics

### 2015-09-10 Core/doc/latex/paper.tex
Authors
paper: one sentencer per line for better version control
paper: is not a thesis

\author[kit]{Thomas M\"uller, Thomas Hauth, Dominik Haitz}
\address[kit]{Karlsruhe Institute of Technology, Wolfgang-Gaede-Stra\ss{}e 1, D-76131 Karlsruhe}

@@

\begin{figure}[p]
\centering \input{figures/analysis_steps}
\caption[Characteristics of an event-based analysis in high energy physics]{Characteristic of an event-based analysis in high energy physics. I/O intensive operations are done in highly parallel in the WLCG. The computing-intensive operations are only performed if really necessary. This means that selection steps are performed as early as possible.}
\label{figure_analysis_steps}
\end{figure}
\clearpage
\section{Requirements and Challenges for Analysis Software\label{section_artus_challenges}}
%define requirements based on previous section, following sections: describe how they are fullfilled
Based on the workflow described in the previous section, the following demands for an analyis framework arise:
\begin{itemize}
 \item Performance
 \begin{itemize}
  \item High performance for fast turnaround cycles, even with large data amounts % -> C++, pipelines, only necessary producers
  \item Flexibility %TODO define need for different configurations in previous section
 \end{itemize}
 \item Maintainability
 \begin{itemize}
  \item Easy to install
  \item Readable and understandable, also for programming newcomers
  \item Compatibility with different operating platforms/environments % standalone compile with make, integrated batch/GC functionality
  \item Compatible with different data formats % -> data format defined outside (kappa)
  \item Clear separation of code, data and settings % -> json config
 \end{itemize}

@@

Currently, the \artus framework is used for various user analysis of CMS data ranging from jet energy calibration and QCD studies to analyses in the scope of Higgs boson measurements and searches for physics beyond the Standard Model. As an example, the technical aspects of the $\PH\to\Pgt\Pgt$ analysis presented in this thesis are explained in the following.

The $\PH\to\Pgt\Pgt$ analysis developed in the scope of this thesis makes use of skimmed data sets in the Kappa format. The Kappa framework\footnote{\url{https://github.com/KappaAnalysis/Kappa}} is designed for skimming CMS data into small n-tuples that can be analysed independently of the CMSSW software. The only dependence is the ROOT framework. The n-tuples store high-level information such as four-momentum vectors, collections of leptons and vertices. Similar as the \artus framework, Kappa aims to satisfy general analysis needs.

@@

The post-processing step either outputs the plots shown in this thesis or writes the histograms to ROOT files and passes them over to the limit calculation tools.

>>>>

\author[kit]{Thomas M\"uller, Thomas Hauth, Dominik Haitz, Raphael Friese}
\address[kit]{Karlsruhe Institute of Technology, Wolfgang-Gaede-Stra\ss{}e 1, D-76131 Karlsruhe}

@@

\begin{figure}[p]
\centering \input{figures/analysis_steps}
\caption[Characteristics of an event-based analysis in high energy physics]{Characteristic of an event-based analysis in high energy physics.
I/O intensive operations are done in highly parallel in the WLCG.
The computing-intensive operations are only performed if really necessary.
This means that selection steps are performed as early as possible.}
\label{figure_analysis_steps}
\end{figure}
\clearpage

\section{Requirements and Challenges for Analysis Software\label{section_artus_challenges}}
%define requirements based on previous section, following sections: describe how they are fullfilled
Based on the workflow described in the previous section, the following demands for an analyis framework arise:
\begin{itemize}
 \item Performance
 \begin{itemize}
  \item High performance for fast turnaround cycles, even with large data amounts % -> C++, pipelines, only necessary producers
  \item Flexibility %TODO define need for different configurations in previous section
 \end{itemize}
 \item Maintainability
 \begin{itemize}
  \item Easy to install
  \item Readable and understandable, also for programming newcomers
  \item Compatibility with different operating platforms/environments % standalone compile with make, integrated batch/GC functionality
  \item Compatible with different data formats % -> data format defined outside (kappa)
  \item Clear separation of code, data and settings % -> json config
 \end{itemize}

@@

Currently, the \artus framework is used for various user analysis of CMS data ranging from jet energy calibration and QCD studies to analyses in the scope of Higgs boson measurements and searches for physics beyond the Standard Model.
As an example, the technical aspects of the $\PH\to\Pgt\Pgt$ analysis are explained in the following.

The $\PH\to\Pgt\Pgt$ analysis makes use of skimmed data sets in the Kappa format.
The Kappa framework\footnote{\url{https://github.com/KappaAnalysis/Kappa}} is designed for skimming CMS data into small n-tuples that can be analysed independently of the CMSSW software.
The only dependence is the ROOT framework.
The n-tuples store high-level information such as four-momentum vectors, collections of leptons and vertices.
Similar as the \artus framework, Kappa aims to satisfy general analysis needs.

@@

The post-processing step either output plots or writes the histograms to ROOT files and passes them over to the limit calculation tools.
